Foreign Exchange Speculation: An Event Study

### 2018-01-15 MDPI/MDPI_template/FinStudies.tex
Added the ReviewComments and changed some references. I have added the changes package to track the changes.
Updated some comments and the FinStudies text with changes.

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
%\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

\begin{document}
\section{Introduction}
If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{KeynesHedge} and \citet{HicksHedge} for example emphasise the role that speculation plays in facilitating financial market transactions. Speculators offer a service for a fee that is determined by the return on their activity.  While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.   However, the later \citet[p. 101]{Keynes1936} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd' and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge. Now speculation is an uninformed and destabilising force.

This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2016 estimated average daily turnover at \$5.1trn with average daily spot transactions of \$1.7trn See \citet{BISFX2016}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}  If speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.

@@

\section{Measuring speculation}
This study uses a unique series of option \emph{risk-reversal skews} to measure of speculative sentiment\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.}, as well as CFTC data on speculative positions. The risk-reversal skew is the relationship between a call option and the equivalent put.    If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%. See \citet{Global} for a technical review. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity strike is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment or related market positioning.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.  See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.

@@

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{FPCum6w}
\caption{Event Study: Extreme (Low $5{th}$ percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme low is at or above the $5^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk. The solid line is the cumulative mean return from the start of the event window; the dashed lines show the 95\% confidence intervals constructed from 1000 random bootstrap samples of appropriate event window day.}
\label{fig:ES3}
\end{figure}

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{FPCum6wa}
\caption{Event Study:  Extreme ($90^{th}$ percentile), Non-commercial net long (S1), 6 week window. Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is at or above the $90^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme reading as a measure of extreme speculative activity.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
\label{fig:ES4}
\end{figure}

@@

\section{Speculation and foreign exchange returns}
If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information about future prices. An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.  However, once extreme, prices are as likely to continue as to reverse. Speculative sentiment and speculative activity influence foreign exchange prices as indicated by the covariance of speculative measures with returns.  These results show that foreign exchange speculation is an important part of the price discovery process.  The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.

@@

It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.  It is less clear whether the results would be reproduced across markets for other financial assets. The results do not rule out uninformed speculation.  The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.  The effect is not evident in this study.  However, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.

This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.  Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.  Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become prominent.  However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.  These results suggest that it could reduce liquidity and increase volatility.

>>>>

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{changes}
%\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

@@

\begin{document}
\section{Introduction}
If speculation is based on information it is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{KeynesHedge} and \citet{HicksHedge} for example emphasise the role that speculation plays in facilitating financial market transactions. Speculators offer a service for a fee that is determined by the return on their activity.  While speculators may be trading on information or noise, informed speculators will tend to drive out the uninformed noise-traders according to \citet{FriedmanPositive}.   However, the later \citet[p. 101]{Keynes1936} regarded speculation as a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd' and \citet{Delong1990noise} provides a model of cases where speculative noise provides a constrain on informed traders benefiting from their knowledge. Now speculation is an uninformed and destabilising force.  \added[remark=1]{We have therefore two concepts of speculation: one that is random and disruptive and one that is informed and beneicial in its provison of liqidity.

The debate about the relative merit of these two concepts of speculation continues to the current era where there has been much discussion over the role that speculation played in the boom and bust of commodities prices in the period between 2006 and 2007.  A common story that has encouraged congressional investigation says that wild speculation drove oil and other commodities prices well above their fundamental value causing harship for commodity consumers and a subsequent collapse that was equally disruptive for the producders.  See \citet{} and \citet{} for fuller details.  However, \citet{} argues that this is not the case and this is backed by \citet{}}

This paper evaluates the relative importance of these two concepts of speculation by investigating the relationship between the intensity of speculation and the weight of speculative positions relative to subsequent movement of foreign exchange prices.\footnote{\label{FX}The foreign exchange market is chosen for this study because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2016 estimated average daily turnover at \$5.1trn with average daily spot transactions of \$1.7trn See \citet{BISFX2016}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).}  \added[remark=2]{The intenisty of speculation is measured with the use of option risk-reversal skew.  As these become more extreme it signals that expectations of buyers and sellers in the foreign exchange options market have diverged significantly from the forward rate.  Weight of speculative positions is measured by the net long or short of speculative accounts in the futures market. If the movement of prices towards equilibrium, \emph{price discovery} is facilitated by order flow from informed traders (as has been asserted theoetically by \citet{Glosten1985Bid} and \citet{Kyle1985Continuous} while being demonstrted empirically by \citet{Lyons1995Microstructure} and \citet{Evans2002Order} and \citet{rime2010exchange}, speculative sentiment and positions would be part of the price discovery process if informed and would disript this process if not. Therefore,}if speculators are informed their activity drives exchange rates towards equilibrium and prices should follow a random walk after these extremes \added[remark=3]{of speculative sentiment or position}; if speculators are uninformed and their activity can be regarded as noise, the most extreme speculative sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place.

@@

\replaced{More recently the debate has move towards the way that noise-trading speculators may affect welfare and how that could be contained. \citet{Tobin1982} and \citet{Tobvin} developed the argument for the eponymous transaction tax that would seek to limit the influence of exchange rate speculation.  \citet{ERW} use a series of fundamentals to show how speculation casued unwarranted devaluations during the period of the European Exchange Rate Mechanism (ERM).   \citet{Tobin3} use this experience to re-state the call for a tax to reduce speculation.  However, if speculation is informed and part of the process of price discovery this tax can be counterproductive as it will reduce liquidity and may cause evengreater volatility. This is a point made by \citet{Mayer} and \citet{McKinnon}.

The call for a tax on speculative activity in financial markets arose again in the middle of the first decade of the new century as a result of surge in commodities prices that was experienced in the period 2005 to 2007. Focused mostly on oil but encompassing many other energy, metal and food commodities, one argument attributed the rise in prices again to speculative actvity.  See..}.

\section{Measuring speculation}
\replaced{This study uses a unique series of option}{A set of} \emph{risk-reversal skews}\added{are used} to measure \delete{of} speculative sentiment\footnote{\deleted{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.}  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.}\delete{, as well as CFTC data on speculative positions.} The risk-reversal skew is the relationship between a call option and the equivalent put.    If the exchange rate is expected to follow a random walk with drift towards the 1 month forward \added[remark=4]{the same maturity as the option}, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a divergence from the random walk with drift is implied in the market price of options.

\added{It is common practice to back market expectations out of information drawn from option prices.  \footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\%. See \citet{Global} for a technical review. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity strike is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment or related market positioning.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, this movement should be reversed.  See \citet{Yan2011} and \citet{FengZhangFriesen} for some analysis of implied volatility in the equity market.

@@

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{FPCum6w}
\caption{Event Study:  Extreme ($90^{th}$ percentile), Non-commercial net long (S1), 6 week window. Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is at or above the $90^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme reading as a measure of extreme speculative activity.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{FPCum6wa}
\caption{Event Study:  Extreme ($90^{th}$ percentile), Non-commercial net long (S1), 6 week window. Non-commercial net per OI (S2) are the net speculative positions (as reported to US regulator the CFTC) relative to all open positions; extreme high is at or above the $90^{th}$ percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme reading as a measure of extreme speculative activity.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}

\label{fig:ES4}
\end{figure}

@@

\section{Speculation and foreign exchange returns}
If foreign exchange speculation is uninformed noise, extreme speculation, whether measured by sentiment or weight of activity, will coincide with deviations from fundamental value; if speculation is informed, speculation is part of the process of price discovery and extremes will provide no information about future prices. An event study showed that speculation is associated with price movement: exchange rates move in the direction of speculative sentiment and activity.  However, once extreme \added{speculative sentiment or extreme speculative positions are established}, prices are as likely to continue as to reverse. Speculative sentiment and speculative activity influence foreign exchange prices as indicated by the covariance of speculative measures with returns.  These results show that foreign exchange speculation is an important part of the price discovery process.  The random walk after the extreme is indicative of foreign exchange prices absorbing most of the information that is available.

@@

It is encouraging that two very different measures of speculative activity produce very similar results and that the results hold over different time frames: one week for the option data and one week for the regulatory position data.  It is less clear whether the results would be reproduced across markets for other financial assets. \added{The foreign exchange market has a particular microstructure that is dominated by over-the-counter transactions in the spot market and this may influence the results.  Further reseach might look at risk-reversals in financial markets that are traded on electronic exchanges to see if these results are replicated.  The data on speculative positions are from the excahnge traded futures market.} \deleted{The results do not rule out uninformed speculation.  The Friedman effect from the Noise-trader model, whereby the informed speculators take advantage of the space created by price pressure from the uniformed, depends on there being \emph{noise-trader risk} or some limits to arbitrage.  The effect is not evident in this study.}  However, the foreign exchange market is a particularly large and liquid market, other asset classes may have alternative institutional features that mean a different relationship or balance between the informed and uninformed and therefore different results.

\added{It is possible to dispute the link between risk reversal and market sentiment. It is not only the direct change in expectations about the future that affect  risk-reversals. The weight of positions from hedging activity could be influential. Therefore future research might also investigate the link between these figures and other measures of short-term sentiment if they exist.}
\deleted{This is important because there is no unambiguous view of speculative activity and because speculation is an important issues in a number of policy debates.  Most notably, the \emph{Tobin Tax} would put a charge against financial transactions, particularly in foreign exchange.  Part of the argument depends on speculation being uninformed as the reduced activity that is caused by the tax would, it is argued, limit volatility and allow more long-term fundamental evaluations to become prominent.  However, if speculation is informed and part of the price discovery process, this tax may have the opposite effect to what is envisaged.  These results suggest that it could reduce liquidity and increase volatility.}
